Active Learning from Blackbox to Timed Connectors

### 2016-04-04 shortpaper/paper.tex
.

% finally no todos should exist in this draft
\usepackage[textsize=tiny]{todonotes}

@@

As a real-time extension of \emph{Reo}, \emph{Timed
Reo}\cite{DBLP:conf/sefm/ArbabBBR04,DBLP:conf/tase/Meng12} is used in this paper to formalize
connectors in DRES. 
Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}, which is a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}. What's more, 
Formal verification and validation techniques have also been proved applicable for timed connectors, 
e.g. a comformance testing framework on timed connectors has been proposed in \cite{DBLP:conf/tase/LiCWS15}.
A SAT-based bounded model checking approach was adapted in \cite{DBLP:journals/scp/Kemper12} for
verifying connectors.

However, most of these tools heavily rely on manually modelling.
\emph{Correctness} of connectors is closely related to some low-level implementation details.
For example, well-written code may behave dramatically weird with an improper set of concurrency
primitives. Such scenarios happen frequently in embedded systems with different hardware platforms or
operating systems. Consequently, manually extracting a proper model from an existing connector
implementation seems rather unreliable, even with reference to its source code.

In this paper we address this question by means of active learning \cite{de2010grammatical,
DBLP:journals/iandc/Angluin87, DBLP:conf/fase/RaffeltS06}.
First of all, we introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo, which can be transformed to concrete Mealy machines with a
specified alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

The rest of the paper is organized as follows: In Section
Section \ref{sec:semantics}, we defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach. 
Finally, Section VI concludes the paper. Due to the limit of space, many technical details are
omitted in this paper.
Please refer to \todo{github repo} for further contents.

@@

In this section, we illustrate how to transform a timed connector into a Mealy machine with timed
action.
Since time is not involved in original Mealy machine, we first discuss how to formalize the time
dimension in Mealy machine. After that, we present the notion of \emph{parameterized Mealy
machine} as a bridge between connectors and Mealy machines.

\todo{time domain?}

@@

\subsection{Parameterized Mealy Machine}
We present a model named \emph{parameterized Mealy machine} to model timed connectors. PMM is
supposed to behave as a bridge between Mealy machines and timed connectors. Connectors are firstly
defined as PMMs, and composed by \emph{product} and \emph{link} operators. Then, concrete
Mealy-machine model can be generated from the PMM model.
Following the formal definition of Mealy machine in Section \ref{sec:preliminaries}, we define the
\emph{parameterized Mealy machine} as follows.

@@

In the definition above, \emph{Input} and \emph{Output} are used to generate the set of input actions
and output actions from the corresponding alphabets and source/sink ends. $Input(\Sigma,I,O)$ is
defined as a set of functions on $I\cup O$ and an additional \emph{time action} T,  
where for all $f\in Input(\Sigma,I,O)\backslash \{T\}$ we have
\[
\forall i\in I.f(i)\in \Sigma\cup\{\bot\}\land \forall o\in O.f(o)\in\{\rread, \rnoread\}
\]
where we use $\bot$ to indicate that there is \emph{no} data item on a channel end. 
Besides, $\rread$ means that a sink end is ready for writing, $\rnoread$ otherwise. 
In the same way,
$Output(\Sigma,O)$ is defined as a set of functions on $O$, with an additional symbol
$\rblock$ indicating an input failure. Similarly, for all $f\in
Output(\Sigma,O)\backslash\{\rblock\}$,
\[
\forall o\in O. f(o)\in \Sigma\cup\{\bot\}
\]

A non-timed input action $a\in Input(\Sigma,I,O)$ can be restricted to $I'\subseteq I$, denoted by
$a\downarrow_{I'}$ which is defined on $Input(\Sigma,I',O)$ and satisfies
\[
\forall i\in I'. a\downarrow_{I'}(i)=a(i)
\]
Restriction on output actions can be defined similarly.

\begin{example}[Input and Output]
  If we have a simple alphabet with only one item $d_0$, a source-end named $A$, and a
  sink-end named $B$, the input actions would be
  \begin{eqnarray*}
    & & Input(\{d_0\},\{A\},\{B\}) \\
    & = & \{\{A\mapsto d_0,B\mapsto\rread\},\{A\mapsto d_0,B\mapsto\rnoread\}, \\
    & & \{A\mapsto\bot,B\mapsto\rread\}, \{A\mapsto\bot,B\mapsto\rnoread\},T\}
  \end{eqnarray*}
  and its output actions
  \[
  Output(\{d_0\},\{A\},\{B\}) =\{\{B\mapsto d_0\},\{B\mapsto \bot\}, \rblock\}
  \]
\end{example}

We use $\varnothing$ to denote the empty output when no data item is written to any sink end.
For example, in this case $\{B\mapsto\bot\}$ can be briefly denoted as $\rempty$.

Parameterized Mealy machines can be seen as an abstraction of Mealy machines, and we
still need a convert function between the two models.

@@

Now we can use PMMs to specify timed Reo. Here we take the asynchronous channel (FIFO1), the
synchronous channel (Sync) and the timed channel (Timer) as examples. For simplicity, we use $A=\_$
to indicate that there is no constraint on channel end $A$.

%\begin{example}[FIFO1]
  %\label{example:pmmfifo}
  %The PMM of a FIFO1 with source end $A$ and sink end $B$ can be defined as follows.
  %\begin{itemize}
    %\item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    %\item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    %\item[-] output function
      %\begin{small}
        %\begin{displaymath}
          %\delta(\Sigma)(s,i)=\left\{
          %\begin{array}[ht]{ll}
            %\rempty & s=q_0\land i=\{A\mapsto\_,B\mapsto\rnoread\} \\
            %\rblock & s=q_0\land i=\{A\mapsto d',B\mapsto\rread\} \\
            %\rblock & s=q_0\land i=\{A\mapsto\bot,B\mapsto\rread\} \\
            %\rempty & s=q_d\land i=\{A\mapsto\bot,B\mapsto\rnoread\} \\     
            %\{B\mapsto d\} & s=q_d\land i=\{A\mapsto\_,B\mapsto\rread\}\\
            %\rblock & s=q_d\land i=\{A\mapsto d',B\mapsto \rnoread\} \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
    %\item[-] transition function
      %\begin{small}
        %\begin{displaymath}
          %\lambda(\Sigma)(s,i)=\left\{
          %\begin{array}[ht]{ll}
            %q_{d'} & s=q_0\land i=\{A\mapsto d',B\mapsto\_\} \\
            %q_{d'} & s=q_{d}\land i=\{A\mapsto d',B\mapsto \rread\} \\
            %q_{d} & s=q_{d}\land i=\{A\mapsto \_,B\mapsto \rnoread\} \\
            %q_0 & otherwise \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
  %\end{itemize}
%\end{example}

%Besides, the concrete Mealy machine (where $\Sigma=\{d\}$) is shown in Fig. \ref{fig:pmmfifo},
%where labels of edges are in form of $\frac{input}{output}$. All $\rblock$ edges are ignored for
%clearance.
%\begin{figure}[ht]
  %\begin{center}
    %\input{./images/pmmfifo.tex}
  %\end{center}
  %\caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{d\}$}
  %\label{fig:pmmfifo}
%\end{figure}

%\begin{example}[Sync]
  %The PMM of a Sync with a source-end A and a sink-end B can be defined as:
  %\begin{itemize}
    %\item[-] $S(\Sigma)=\{q_0\}$, $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    %\item[-] output function
      %\begin{displaymath}
        %\delta(\Sigma)(s,i)=\left\{
        %\begin{array}[ht]{ll}
          %\rempty & i=\{A\mapsto\bot,B\mapsto\rnoread\} \\
          %\{B\mapsto d\} & i=\{A\mapsto d,B\mapsto\rread\}\\
          %\rblock & i=\{A\mapsto\bot,B\mapsto\rread\} \\
          %\rblock & i=\{A\mapsto d,B\mapsto\rnoread\} \\
        %\end{array}
        %\right.
      %\end{displaymath}
    %\item[-] transition function $\lambda(\Sigma)(s,i)=q_0$.
  %\end{itemize}
%\end{example}

@@

\begin{definition}[Product]
  For two PMMs $\mathcal{PM}_1$ and $\mathcal{PM}_2$, if
  $\mathcal{PM}_2.O\cap\mathcal{PM}_1.I=\varnothing$, their product 
  \[
  \mathcal{PM'} = prod(\mathcal{PM}_1,\mathcal{PM}_2)  \]
  can be defined as follows.
  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}_1.S(\Sigma)\times \mathcal{PM}_2.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=(\mathcal{PM}_1.I\cup \mathcal{PM}_2.I)\backslash \mathcal{PM}_1.O$
    \item[-] $\mathcal{PM}'.O=(\mathcal{PM}_1.O\cup \mathcal{PM}_2.O)\backslash \mathcal{PM}_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $\mathcal{PM}_1$ can be connected to source ends of
  $\mathcal{PM}_2$, but not vise versa)}
  \begin{itemize}
    \item[-] $\mathcal{PM}'.s_0=(\mathcal{PM}_1.s_0, \mathcal{PM}_2.s_0)$
    \item[-] $\forall\Sigma. \mathcal{PM}'.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (\bigcup_{j=1,2} Out_j)\downarrow_{\mathcal{PM}'.O} & i\neq T\land(\bigwedge_{j=1,2} Out_j\neq\rblock) \\
          \rempty & i = T \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $i\in Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O)$
        \item[*] $In_1 = i\downarrow_{\mathcal{PM}_1.I}$
        \item[*] $In_2 = (Out_1 \cup i)\downarrow_{\mathcal{PM}_2.I}$
        \item[*] $Out_j = \mathcal{PM}_j.\delta(\Sigma)(s_j,In_j)$ for $j=1,2$
      \end{itemize}
    \item[-] $\forall\Sigma. \mathcal{PM}'. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where for $j=1,2$, $s_j' = \mathcal{PM}_j.\lambda(\Sigma)(s_j,In_j)$.
  \end{itemize}
\end{definition}

The idea in \emph{product} is quite simple. The output of $\mathcal{PM}_1$ would be provided as part of the
input of $\mathcal{PM}_2$. Timed action $T$ is only executed simultaneously between $\mathcal{PM}_1$
and $\mathcal{PM}_2$.

As mentioned above, we cannot connect a sink end of $\mathcal{PM}_2$ to a source end of
$\mathcal{PM}_1$, which makes it difficult to construct connectors like alternator in Fig.
\ref{fig:reoconnector}. To solve the problem, we define a \emph{link} operator to connect sink
ends and source ends in the same connector.

\begin{definition}[Link]
  The \emph{link} operator constructs connector by connecting a sink end $\OUT$ to a source end $\IN$
  within the same connector.   
  \[
  \mathcal{PM}' = link(\mathcal{PM}, \OUT, \IN)
  \]
  Here we assume that $\IN\in \mathcal{PM}.I$ and $\OUT\in \mathcal{PM}.O$.

  \begin{itemize}
  	\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}.S(\Sigma)$
    \item[-] $\mathcal{PM}'.I=\mathcal{PM}.I\backslash\{\IN\}$
    \item[-] $\mathcal{PM}'.O=\mathcal{PM}.O\backslash\{\OUT\}$
    \item[-] $\mathcal{PM}'.s_0=\mathcal{PM}.s_0$
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).\mathcal{PM}'.\delta(\Sigma)(s,i)=$
      \begin{small}
        \begin{displaymath}
          \left\{
          \begin{array}[h]{lr}
            \rempty & i=T \\
            \mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})\downarrow_{\mathcal{PM}'.O} & (*)\\
            \rblock & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
      where the condition (*) is defined as
      \[
      \exists d.\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})(\OUT)=d
      \]
    \item[-] $\forall\Sigma, i\in
      Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).$
      \begin{small}
        \begin{displaymath}
          \mathcal{PM}.\lambda(\Sigma)(s,i)=
          \left\{
          \begin{array}[ht]{lr}
            \mathcal{PM}.\lambda(\Sigma,T) & i=T \\
            \mathcal{PM}.\lambda(\Sigma,i\cup\{\IN\mapsto d\}) & (*)\\
            s & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
  \end{itemize}
\end{definition}

@@

\subsection{Observation Table}
\emph{Observation Tables}, proposed in \cite{DBLP:journals/iandc/Angluin87}, are used to construct
hypothetical Mealy machines from the results of membership queries.

Mealy machines consist of \emph{states} and \emph{transitions}.
As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
state of $s$ after executing $a$.
However, under such notations, there are \emph{infinite} number of states, among which most are
duplicate.

States are distinguished, usually, by their \emph{subsequent behavior}. Suppose $s_1\neq s_2\in
\mathcal{I}^*$, we say the corresponding of states $s_1$ and $s_2$ are equivalent iff $\forall d\in
\mathcal{I}^+. mq(s_1d)=mq(s_2d)$. Since we can never check it on every $d\in\mathcal{I}^+$, an
alternative way is to use a set of \emph{suffixes} to distinguish them approximately.
\begin{definition}[$D$-Equivalence]
  Provided with a set of suffixes $D\subset\mathcal{I}^+$, the corresponding states of
  $s_1,s_2\in\mathcal{I}^*$ are $D$-equivalent (denoted as $s_1\sim_D s_2$)
  iff $mq(s_1d) = mq(s_2d)$ for all $d\in D$.
\end{definition}

Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
sequence $acc(s,D)$ of $s$ is defined as the shortest sequence in $\{s'\in\mathcal{I}^*|s'\sim_D
s\}$\footnote{When there are multiple satisfying sequences with the same length, we use a certain tactic
to pick up one from them determinstically.}.

Formally, an observation table (see Fig. \ref{fig:hypo} as an example) is determined by a tuple $(S,D)$ where
$S\subset\mathcal{I}^*$ is a set of the access sequences of states and $D\subset\mathcal{I}^*$
is a set of suffixes. In such tables, columns are labelled with suffixes in $D$, and rows are
labelled with input sequences. As shown in Fig. \ref{fig:hypo}, rows are divided into two parts. The upper
part, labelled with access sequence $s\in S$, denotes the states covered in the current hypothesis.
The lower part, labelled with sequences in $\{sa|s\in S,a\in\mathcal{I}\}\backslash S$ (representing
any potentially unclosed transition) denotes all the transitions outside $s\in S$. A cell with row
label $s$ and column label $d$ is filled with $mq(sd)$.

An observation table $(S,D)$ is \emph{closed} if 
\[
\forall s\in S,a\in\mathcal{I}.\exists s'\in S. sa\sim_D s'
\]

From a closed observation table $obs$, we can construct a hypothetical Mealy machine $\mathcal{H}=\langle
S,s_0,\mathcal{I},\mathcal{O},\delta,\lambda\rangle$ in a straight-forward way:
\begin{itemize}
  \item[-] Every state in $S$ corresponds to an input sequence in $obs.S$
  \item[-] $s_0$ corresponds to the empty sequence $\varepsilon$

    \emph{(we use input sequences to indicate their corresponding states in following parts)}
  \item[-] $\delta(s,a)=mq(sa)$
  \item[-] $\lambda(s,a)=acc(sa,obs.D)$. Since $obs$ is closed, we always have $\lambda(s,a)\in S$.
\end{itemize}

Algorithm \ref{alg:buildtable} shows how to build a closed observation table.

\begin{algorithm} 
  \caption{EncloseTable} 
  \label{alg:buildtable}
  \small
  \KwIn{An oracle interface $mq$, An input actions $\mathcal{I}$, An observation table $obs$}
  \KwOut{A closed observation table $obs'$} 
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in obs.S,\forall t\in \mathcal{I}\}\backslash obs.S$\;
    $unclosed=\{seq\in next|\forall a\in obs.S, seq \not\sim_{obs.D} a\}$\;
    $obs'.S$ = $obs.S\cup unclosed$\;
    $obs'.D$ = $obs.D$\;
  }
  \Return $obs'$\; 
\end{algorithm}

Taking a 2-Timer channel as an example, we now illustrate how this algorithm works. 
We assume that the source end of this 2-Timer channel is $A$, the sink end is $B$, and the alphabet
is $\{a\}$. We briefly denote $\{A\mapsto a,B\mapsto \rnoread\}$ and other input/output actions in form of
$(a,\rnoread)$.

Firstly, $obs.S$ is initialized with the initial state (denoted by $\varepsilon$), and $obs.D$ is
initialized with all one-step suffixes.
Then we explore all the access sequences in $obs.S$ and calculate its successors. Here
$unclosed$ consists of access sequences that has no equivalent sequence in $obs.S$. 
For example, Fig. \ref{fig:hypo} shows the $obs$ of a 2-Timer channel after the first
iteration. The five successors of $\varepsilon$ are presented at the bottom of the table, where
four are equivalent with $\varepsilon$ but $(a,\rnoread)$ is not. Therefore, we take $(a,\rnoread)$ as a
brand-new state and all of its successors need further exploration.

>>>>

% finally no todos should exist in this draft
\usepackage[textsize=tiny]{todonotes}
\usepackage{xcolor}
\newcommand{\liyi}[1]{\textcolor{red}{Liyi: {#1}}}

@@

As a real-time extension of \emph{Reo}, \emph{Timed
Reo}\cite{DBLP:conf/sefm/ArbabBBR04,DBLP:conf/tase/Meng12} is used in this paper to formalize
connectors in DRES. 
Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}, which is a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}. 
\begin{figure}[ht]
  \begin{center}
    \input{./images/basic.tex}
  \end{center}
  \caption{Basic Reo Channels}
  \label{fig:basic}
\end{figure}

Formal verification and validation techniques have also been proved applicable for timed connectors
\cite{DBLP:conf/tase/LiCWS15,DBLP:journals/scp/Kemper12} 
However, most of these tools heavily rely on manually modelling.
\emph{Correctness} of connectors is closely related to some low-level implementation details.
For example, well-written code may behave dramatically weird with an improper set of concurrency
primitives. Such scenarios happen frequently in embedded systems with different hardware platforms or
operating systems. Consequently, manually extracting a proper model from an existing connector
implementation seems rather unreliable, even with reference to its source code.

In this paper we address this question by means of active learning
\cite{DBLP:journals/iandc/Angluin87, DBLP:conf/fase/RaffeltS06}.
First of all, we introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo, which can be transformed to concrete Mealy machines with a
specified alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

The rest of the paper is organized as follows: In Section \ref{sec:semantics}, we defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach. 
Finally, Section VI concludes the paper. Due to the limit of space, many technical details are
omitted in this paper.
Please refer to \cite{reo-learn-github} for further contents.

@@

In this section, we introduce PMM as an intermediate form to transform timed connectors into
Mealy machines with timed action. In this paper, we assume that time dimension is defined on the
national number $\mathbb{Q}$, which is generally as expressive as $\mathbb{R}$ and much easier to
formalize\cite{DBLP:journals/fmsd/PrabhakarDM015}. Since connectors contain only finite number of
channels, we can always find a minimal time unit (denoted as $T$) where real-time connectors can be
discretized.

@@

\subsection{Parameterized Mealy Machine}
In this work, we use \emph{Mealy machines} to model Reo connectors.
Mealy machine is an extension of finite state machine to model reactive systems.
Following its definition in \cite{DBLP:conf/sfm/SteffenHM11}, we present a model here named
\emph{parameterized Mealy machine} to formalize timed connectors.

@@

In the definition above, \emph{Input} and \emph{Output} are used to generate the set of input actions
and output actions from the corresponding alphabets and source/sink ends. $Input(\Sigma,I,O)$ is
defined as the intersection of the set of input events $Evt_{in}$ and an additional \emph{time
action} T. Here every $in\in Evt_{in}$ is an evaluation on both sink and source ends and
\[
\forall i\in I.in(i)\in\Sigma\cup\{\rempty\}\land \forall o\in O.in(o)\in\{\rnoread, \rread\}
\]
where we use $\bot$ to indicate that there is \emph{no} data item on a channel end. 
Besides, $\rread$ means that a sink end is ready for writing, $\rnoread$ otherwise. 
In the same way,
$Output(\Sigma,O)$ is defined as a set of output events in form of evaluations on $O$, with an
additional symbol $\rblock$ indicating an input failure.

%A non-timed input action $a\in Input(\Sigma,I,O)$ can be restricted to $I'\subseteq I$, denoted by
%$a\downarrow_{I'}$ which is defined on $Input(\Sigma,I',O)$ and satisfies
%\[
%\forall i\in I'. a\downarrow_{I'}(i)=a(i)
%\]
%Restriction on output actions can be defined similarly.

\begin{example}[Input and Output]
  If we have a simple alphabet with only one item $d_0$, a source-end named $A$, and a
  sink-end named $B$, the input actions would be
  \begin{eqnarray*}
    & & Input(\{d_0\},\{A\},\{B\}) \\
    & = & \{\{A\mapsto d_0,B\mapsto\rread\},\{A\mapsto d_0,B\mapsto\rnoread\}, \\
    & & \{A\mapsto\bot,B\mapsto\rread\}, \{A\mapsto\bot,B\mapsto\rnoread\},T\}
  \end{eqnarray*}
  and its output actions
  \[
  Output(\{d_0\},\{A\},\{B\}) =\{\{B\mapsto d_0\},\varnothing, \rblock\}
  \]
\end{example}

We use $\varnothing$ to denote the empty output when no data item is written to any sink end.
For example, in the above case $\varnothing$ is used to denote $\{B\mapsto\bot\}$.

Parameterized Mealy machines can be seen as an abstraction of Mealy machines, hence we
still need a convert function between the two models.

@@

Now we can use PMMs to specify timed Reo. Here we take the timed channel (Timer) as an example. In
timed Reo, a n-timer accepts a data item from its source end, and then put it to its sink end after
n time units. Throughout the duration, the channel cannot accept any more data items.
For simplicity, we use $A=\_$ to indicate that there is no constraint on channel end $A$.

@@

\liyi{Actually I'm not sure if omitting definition of prod and link is a good idea. Generally I
think this paper should focus on two parts: definition of PMM and optimization of L* algorithm.}

%\begin{definition}[Product]
  %For two PMMs $\mathcal{PM}_1$ and $\mathcal{PM}_2$, if
  %$\mathcal{PM}_2.O\cap\mathcal{PM}_1.I=\varnothing$, their product 
  %\[
  %\mathcal{PM'} = prod(\mathcal{PM}_1,\mathcal{PM}_2)  \]
  %can be defined as follows.
  %\begin{itemize}
    %\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}_1.S(\Sigma)\times \mathcal{PM}_2.S(\Sigma)$
    %\item[-] $\mathcal{PM}'.I=(\mathcal{PM}_1.I\cup \mathcal{PM}_2.I)\backslash \mathcal{PM}_1.O$
    %\item[-] $\mathcal{PM}'.O=(\mathcal{PM}_1.O\cup \mathcal{PM}_2.O)\backslash \mathcal{PM}_2.I$
  %\end{itemize}
  %\emph{(Here we assume that sink ends of $\mathcal{PM}_1$ can be connected to source ends of
  %$\mathcal{PM}_2$, but not vise versa)}
  %\begin{itemize}
    %\item[-] $\mathcal{PM}'.s_0=(\mathcal{PM}_1.s_0, \mathcal{PM}_2.s_0)$
    %\item[-] $\forall\Sigma. \mathcal{PM}'.\delta(\Sigma)((s_1,s_2), i)=$
      %\begin{displaymath}
        %\left\{
        %\begin{array}[ht]{ll}
          %(\bigcup_{j=1,2} Out_j)\downarrow_{\mathcal{PM}'.O} & i\neq T\land(\bigwedge_{j=1,2} Out_j\neq\rblock) \\
          %\rempty & i = T \\
          %\rblock & otherwise \\
        %\end{array}
        %\right.
      %\end{displaymath}
      %where we have
      %\begin{itemize}
        %\item[*] $i\in Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O)$
        %\item[*] $In_1 = i\downarrow_{\mathcal{PM}_1.I}$
        %\item[*] $In_2 = (Out_1 \cup i)\downarrow_{\mathcal{PM}_2.I}$
        %\item[*] $Out_j = \mathcal{PM}_j.\delta(\Sigma)(s_j,In_j)$ for $j=1,2$
      %\end{itemize}
    %\item[-] $\forall\Sigma. \mathcal{PM}'. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      %where for $j=1,2$, $s_j' = \mathcal{PM}_j.\lambda(\Sigma)(s_j,In_j)$.
  %\end{itemize}
%\end{definition}

%The idea in \emph{product} is quite simple. The output of $\mathcal{PM}_1$ would be provided as part of the
%input of $\mathcal{PM}_2$. Timed action $T$ is only executed simultaneously between $\mathcal{PM}_1$
%and $\mathcal{PM}_2$.

%As mentioned above, we cannot connect a sink end of $\mathcal{PM}_2$ to a source end of
%$\mathcal{PM}_1$, which makes it difficult to construct connectors like alternator in Fig.
%\ref{fig:reoconnector}. To solve the problem, we define a \emph{link} operator to connect sink
%ends and source ends in the same connector.

%\begin{definition}[Link]
  %The \emph{link} operator constructs connector by connecting a sink end $\OUT$ to a source end $\IN$
  %within the same connector.   
  %\[
  %\mathcal{PM}' = link(\mathcal{PM}, \OUT, \IN)
  %\]
  %Here we assume that $\IN\in \mathcal{PM}.I$ and $\OUT\in \mathcal{PM}.O$.

  %\begin{itemize}
    %\item[-] $\forall\Sigma. \mathcal{PM}'.S(\Sigma)=\mathcal{PM}.S(\Sigma)$
    %\item[-] $\mathcal{PM}'.I=\mathcal{PM}.I\backslash\{\IN\}$
    %\item[-] $\mathcal{PM}'.O=\mathcal{PM}.O\backslash\{\OUT\}$
    %\item[-] $\mathcal{PM}'.s_0=\mathcal{PM}.s_0$
    %\item[-] $\forall\Sigma, i\in
      %Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).\mathcal{PM}'.\delta(\Sigma)(s,i)=$
      %\begin{small}
        %\begin{displaymath}
          %\left\{
          %\begin{array}[h]{lr}
            %\rempty & i=T \\
            %\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})\downarrow_{\mathcal{PM}'.O} & (*)\\
            %\rblock & otherwise \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
      %where the condition (*) is defined as
      %\[
      %\exists d.\mathcal{PM}.\delta(\Sigma)(s,i\cup\{\IN\mapsto d\})(\OUT)=d
      %\]
    %\item[-] $\forall\Sigma, i\in
      %Input(\Sigma,\mathcal{PM}'.I,\mathcal{PM}'.O).$
      %\begin{small}
        %\begin{displaymath}
          %\mathcal{PM}.\lambda(\Sigma)(s,i)=
          %\left\{
          %\begin{array}[ht]{lr}
            %\mathcal{PM}.\lambda(\Sigma,T) & i=T \\
            %\mathcal{PM}.\lambda(\Sigma,i\cup\{\IN\mapsto d\}) & (*)\\
            %s & otherwise \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
  %\end{itemize}
%\end{definition}

@@

\subsection{Active Learning: Overview}
In this section, we briefly introduce the main ideas of active automata learning.

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
under certain inputs. By well-designed query strategy, active learning is able to obtain more
accurate models with smaller data-set.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.5\textwidth]{./images/activelearning.png}
  \end{center}
  \caption{Active Automata Learning}
  \label{fig:activelearning}
\end{figure}

Fig. \ref{fig:activelearning} shows the sketch of active learning, wherein:
\begin{itemize}
  \item[-] \emph{Teacher} and \emph{Student}: Active learning is an interactive process where
    students ask questions and teachers answer. Here learning algorithm plays the role of student.
  \item[-] \emph{Oracles} is an interface specifying which kind of questions can be answered by the
    teacher.
  \item[-] \emph{SUL} is an abbreviation of System Under Learn. In this paper, we take blackbox
    models as our SULs.
  \item[-] \emph{CEGAR} indicates Counter-Example Guided Abstraction
    Refinement\cite{DBLP:conf/cav/ClarkeGJLV00}. In active learning, we need counter-examples to
    guide us on further queries and cover the undistinguished states.
\end{itemize}

When applying active automata learning on some model, we assume that the model should be equivalent
to some \emph{Mealy machine}, i.e., a deterministic model accepting a finite set of inputs and
mapping them to a finite set of outputs.

Such a model is encapsulated as a \emph{teacher} by the \emph{Oracle}, which handles
all communication with the model. The oracle also serves as a so-called \emph{Minimal Adequate Teacher}
interface, which is responsible for two types of queries.

\begin{itemize}
  \item[-] \textbf{Membership Query} (hereinafter referred to as \emph{mq}) In grammar-learning
    \cite{DBLP:journals/iandc/Angluin87}, \emph{mq} checks if a word is a member of certain language
    defined by the given grammar. When it comes to
    automata learning, \emph{mq} is supposed to provide \emph{simulation results} for given input
    sequences.
  \item[-] \textbf{Equivalence Query} (hereinafter referred to as \emph{eq}) Given a hypothesis
    (usually constructed by the learning algorithm), \emph{eq} checks whether the hypothesis is
    equivalent to the system-under-learn and generates a counter-example if needed. Generally,
    \emph{eq} is irrealizable when SUL is a blackbox. So we tactically use $mq$ to achieve the
    approximate results.
\end{itemize}

These queries are given by \emph{learning algorithms}, or so-called \emph{students} in Fig.
\ref{fig:activelearning}. From the \emph{mq} results, a learning algorithm constructs a
\emph{hypothesis} and then check it with \emph{eq}. If counter-examples are found, we turn back and
repeat the hypothesis construction until the equivalence query returns \emph{true}.

More details on the active automata learning algorithm will be presented in Section
\ref{sec:activelearning}.

%\subsection{Observation Table}
%\emph{Observation Tables}, proposed in \cite{DBLP:journals/iandc/Angluin87}, are used to construct
%hypothetical Mealy machines from the results of membership queries.

%Mealy machines consist of \emph{states} and \emph{transitions}.
%As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
%after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
%state of $s$ after executing $a$.
%However, under such notations, there are \emph{infinite} number of states, among which most are
%duplicate.

%States are distinguished, usually, by their \emph{subsequent behavior}. Suppose $s_1\neq s_2\in
%\mathcal{I}^*$, we say the corresponding of states $s_1$ and $s_2$ are equivalent iff $\forall d\in
%\mathcal{I}^+. mq(s_1d)=mq(s_2d)$. Since we can never check it on every $d\in\mathcal{I}^+$, an
%alternative way is to use a set of \emph{suffixes} to distinguish them approximately.
%\begin{definition}[$D$-Equivalence]
  %Provided with a set of suffixes $D\subset\mathcal{I}^+$, the corresponding states of
  %$s_1,s_2\in\mathcal{I}^*$ are $D$-equivalent (denoted as $s_1\sim_D s_2$)
  %iff $mq(s_1d) = mq(s_2d)$ for all $d\in D$.
%\end{definition}

%Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
%sequence $acc(s,D)$ of $s$ is defined as the shortest sequence in $\{s'\in\mathcal{I}^*|s'\sim_D
%s\}$\footnote{When there are multiple satisfying sequences with the same length, we use a certain tactic
%to pick up one from them determinstically.}.

%Formally, an observation table (see Fig. \ref{fig:hypo} as an example) is determined by a tuple $(S,D)$ where
%$S\subset\mathcal{I}^*$ is a set of the access sequences of states and $D\subset\mathcal{I}^*$
%is a set of suffixes. In such tables, columns are labelled with suffixes in $D$, and rows are
%labelled with input sequences. As shown in Fig. \ref{fig:hypo}, rows are divided into two parts. The upper
%part, labelled with access sequence $s\in S$, denotes the states covered in the current hypothesis.
%The lower part, labelled with sequences in $\{sa|s\in S,a\in\mathcal{I}\}\backslash S$ (representing
%any potentially unclosed transition) denotes all the transitions outside $s\in S$. A cell with row
%label $s$ and column label $d$ is filled with $mq(sd)$.

%An observation table $(S,D)$ is \emph{closed} if 
%\[
%\forall s\in S,a\in\mathcal{I}.\exists s'\in S. sa\sim_D s'
%\]
